---
title: "Reseach Projects"
subtitle: "Soc 225: Data & Society"
author: "Raden Roy Pradana"
date: "`r Sys.Date()`"
output: html_document
---

Import the libraries

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(sf)
library(leaflet)
```

Read the crime data

```{r}
# read teh crime data
crime_data <- read_csv("data/Crime_Data.csv")

# make date as a date type
crime_data <- crime_data %>%
  mutate(Date = as.Date(`Reported Date`, "%m/%d/%Y"))

# filter based on date
crime_data <- crime_data %>%
  filter(Date >= as.Date("2018-01-01") & Date < ("2019-01-01"))
         
# rename Beat to beat
crime_data <- crime_data %>%
  rename(beat = Beat)

# drop unknown beat
crime_data <- crime_data %>%
  filter(!is.na(beat))

# dreap unknown neighborhood
crime_data <- crime_data %>%
  filter(Neighborhood != "UNKNOWN")

crime_data
```

```{r}
unique(crime_data$beat)
```

# Upload the school rating from Seattle Public School files

```{r}
school_rating <- read_csv("data/school_rate.csv")

# filter school that doesn't have score
school_rating <- school_rating %>%
  filter(English != 0)

school_rating
```



# Read the shape file for School location

```{r}
school_map <- st_read("data/SPS_AttendanceAreasAndSchools_Shapefiles_2019_2020/sites/sites_2019.shp")

# filter only elemtary, middle school, and High School
school_map <- school_map %>%
  filter(Status == "HS" | Status == "MS" | Status == "ELEM")
school_map
```

# Join School map and its rating

```{r}
# join school rating and its map
school_map_data <- inner_join(school_map, school_rating, by="PROPERTY_L")
school_map_data
```


# Load the police beat shape file for mapping the beat

```{r}
police_beats_map <- st_read("data/Seattle_Police_Beats_2018Present/Seattle_Police_Beats_2018Present.shp")
police_beats_map
```

# Data exploration

```{r}
crime_data %>%
  ggplot(aes(Neighborhood)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_y_log10() +
  geom_bar()
```


```{r}
crime_data %>%
  ggplot(aes(beat)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  geom_bar()
```


# Count the number of crimes on each beat and joining the number of beat and its map location

```{r}
crime_data_group <- crime_data %>%
  group_by(beat) %>%
  count()

crime_data_group <- crime_data_group %>%
  rename(
    numberOfCrime = n
  )

crime_map_data <- inner_join(police_beats_map, crime_data_group, by="beat")
```

# Plot the number of crime on the map

```{r}
ggplot(crime_map_data) +
  geom_sf(aes(fill=numberOfCrime)) + 
  geom_sf(data=school_map_data, size=0.5, color="black") +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Seattle Crime Map") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  scale_fill_gradient(low="white", high="red") +
  coord_sf()

#ggsave("Seattle_Crime_Map.png")
```

# Combining between school map and data, and data from Seattle PD

```{r}
school_to_beat <- st_join(school_map_data %>% st_transform(4326), crime_map_data, join=st_within)
school_to_beat
```

# calculate the average score in for each beat

```{r}
beat_rating <- data.frame(school_to_beat) %>%
  group_by(beat) %>%
  summarise(english_score=mean(English), math_score=mean(Math), avg_score=(english_score + math_score) / 2) %>%
  arrange(desc(avg_score))

beat_rating
```

# Combine the crime map and the school rating for each beat
```{r}
score_map <- inner_join(crime_map_data, beat_rating)
score_map_df <- data.frame(score_map)
```

```{r}
score_map_df
```

# Plot the line graph

```{r}
score_map_df %>%
  ggplot(aes(x=numberOfCrime, y=avg_score)) +
  geom_point() +
  ylim(15,100) +
  ggtitle("Test vs. Number of Crime") +
  geom_smooth()

ggsave("Total Average Score Line.png")


score_map_df %>%
  ggplot(aes(x=numberOfCrime, y=math_score)) +
  ylim(15,100) +
  geom_point() +
  ggtitle("Mathematics Proficiency Test vs. Number of Crime") +
  geom_smooth()

ggsave("Math Profiency Test line.png")


score_map_df %>%
  ggplot(aes(x=numberOfCrime, y=english_score)) +
  ylim(15,100) +
  geom_point() +
  ggtitle("English Proficiency Test vs. Number of Crime") +
  geom_smooth()

ggsave("English Profiency Test line.png")

```

# Plot the map

```{r}
ggplot() +
  geom_sf(data=crime_map_data, size=0.8, color="black") +
  geom_sf(data=score_map,aes(fill=avg_score)) + 
  geom_sf(data=school_map_data, size=1, color="black") +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Total Average Score") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  scale_fill_gradient(low="red", high="blue") +
  coord_sf()

#ggsave("Total Average Score.png")

ggplot() +
  geom_sf(data=crime_map_data, size=0.8, color="black") +
  geom_sf(data=score_map,aes(fill=english_score)) + 
  geom_sf(data=school_map_data, size=1, color="black") +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("English Proficiency Test Score") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  scale_fill_gradient(low="red", high="blue") +
  coord_sf()

#ggsave("English Profiency Test Score.png")

ggplot() +
  geom_sf(data=crime_map_data, size=0.8, color="black") +
  geom_sf(data=score_map,aes(fill=math_score)) + 
  geom_sf(data=school_map_data, size=1, color="black") +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Mathematics Proficiency Test Score") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  scale_fill_gradient(low="red", high="blue") +
  #scale_fill_distiller(palette = "Spectral") +
  coord_sf()

#ggsave("Mathematics Profiency Test Score.png")


ggplot(crime_map_data) +
  geom_sf(aes(fill=numberOfCrime)) + 
  geom_sf(data=school_map_data, size=1, color="black") +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Seattle Crime and School Map") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  scale_fill_gradient(low="white", high="red") +
  coord_sf()
```

# Read census data to find population and crime ratio

```{r}
waPop <- readRDS("data/waPop.Rds")
```

# Filter the census to King County

```{r}
waPop_kc <- waPop %>%
  filter(COUNTYFP10 == "033")
waPop_kc
```

# Combine the census and the police beat map

```{r}
beat_to_pop <- st_join(police_beats_map, waPop_kc %>% st_transform(4326))

# drop the area outside the police beat
pop_per_beat <- beat_to_pop %>%
  filter(!is.na(beat))
```

# Group the data by tracts census block

```{r}
pop_per_beat <- pop_per_beat %>%
  select(Population, TRACTCE10, beat) %>%
  group_by(TRACTCE10)

pop_per_beat
```



```{r}
# Join the population per beat with the school score map by beat
pop_per_beat <- left_join(pop_per_beat, score_map_df, by="beat")

# filter area withhout crime
pop_per_beat <- pop_per_beat %>%
  filter(!is.na(numberOfCrime))
pop_per_beat
```

```{r}
# Create a new dataframe after we group everything by beat, and find its summarise
pop_per_beat_group <- pop_per_beat %>%
  group_by(beat) %>%
  summarise(total_crime = sum(numberOfCrime), 'english score' = mean(english_score), 'math score' = mean(math_score), 'avg score' = mean(avg_score),pop = sum(Population))

pop_per_beat_group
```

# Plot the school rating based on the popuation and total crime ratio

```{r}
# Create the population and crime ratio 
pop_per_beat_group <- pop_per_beat_group %>%
  mutate('ratio population and crime' = pop / total_crime )

# Plot english score
pop_per_beat_group %>%
  ggplot(aes(x=`ratio population and crime`, y=`english score`)) +
  geom_point() +
  ylim(20,100) +
  geom_smooth() +
  ggtitle("English Proficiency Score vs Percentage Crime")

ggsave("English Profiency Test Score new.png")

# Plot Math score
pop_per_beat_group %>%
  ggplot(aes(x=`ratio population and crime`, y=`math score`)) +
  geom_point() +
  ylim(20,100) +
  geom_smooth() +
    ggtitle("Math Proficiency Score vs Percentage Crime")

ggsave("Mathematics Profiency Test Score new.png")

# Plot average score
pop_per_beat_group %>%
  ggplot(aes(x=`ratio population and crime`, y=`avg score`)) +
  geom_point() +
  ylim(20,100) +
  geom_smooth() +
  ggtitle("Overall Proficiency Score vs Percentage Crime")

ggsave("Average Profiency Test Score new.png")
  
```

```{r}
# plot the map using the new population and crime ratio
ggplot() +
  geom_sf(data=crime_map_data, size=0.8, color="black") +
  geom_sf(data=pop_per_beat_group, aes(fill=`ratio population and crime`)) + 
  geom_sf(data=school_map_data, size=1, color="black") +
  xlab("Longitude") + ylab("Latitude") +
  ggtitle("Seattle Crime and School Map") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  scale_fill_gradient(low="white", high="red") +
  coord_sf()

#ggsave("Seattle_Crime_Map_new.png")
```


